---
title: 'Task 3: Santa Barbara urchins'
author: "Anna Zauner"
date: "10/23/2021"
output: html_document
---

# Overview

This task utilizes data from 5 sites along the Santa Barbara coast to ....

**Data citation**: Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b. 


# Set up

```{r setup, include=TRUE, warning= FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# Attach packages
library(tidyverse)
library(dplyr)
library(janitor)
library(here)
library(kableExtra)

```

# Read in data

Read in with `read_csv` and cleaned names with `clean_names()`
```{r}
urchins_data <- read_csv(here("data", "urchins.csv")) %>% 
clean_names() %>% 
rename(size_cm = size)

```

## Section 1

Size comparisons, purple and red urchins. 

### Visualize the data

Created a faceted histogram to compare the distribution of sizes between purple and red urchins. 

```{r}
ggplot(data = urchins_data, aes(x= size_cm))+
  geom_histogram(color= "black", fill= "white", bin= 10)+
  facet_grid(common_name ~ .)+
  labs(x = "Size (cm)",
       y = "Count",
       title = "Distibution of sizes between purple and red urchins")
```
Figure 1: **Distribution of sizes (cm) of purple and red urchins** Comparing the distribution of sizes in centimeters between the purple and red urchins off the coast of Santa Barbara. 

### Descriptive statistics

Created a table to summarize statistic for purple and red urchins by finding mean, median, standard deviation and sample size. 

```{r}
urchins_summary <- urchins_data %>% 
  group_by(common_name) %>% 
  summarize(mean = mean(size_cm),
            median = median(size_cm),
            sd = sd(size_cm)) %>% 
  kable(col.names = c("Urchin name",
                      "Mean (cm)",
                      "Median (cm)",
                      "Standard Deviation"),
        digits = 1)

urchins_summary
```


Takeaways: The purple urchin population has a lower central tendency (4.5cm mean, 4.5cm median) than the red urchin population (6.8cm mean, 7.0cm median). The purple urchins are more concentrated around the central tendency, with a smaller spread and lower standard deviation (1.2) while red urchins are more spread with a higher standard deviation (1.9).