---
title: 'Task 3: Santa Barbara urchins'
author: "Anna Zauner"
date: "10/23/2021"
output: html_document
---

# Overview

This task utilizes data from 5 sites along the Santa Barbara coast to ....

**Data citation**: Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b. 


# Set up

```{r setup, include=TRUE, warning= FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# Attach packages
library(tidyverse)
library(dplyr)
library(janitor)
library(here)


```

# Read in data

Read in with `read_csv` and cleaned names with `clean_names()`
```{r}
urchins_data <- read_csv(here("data", "urchins.csv"))

clean_names(urchins_data)

```

## Section 1

Size comparisons, purple and red urchins. 

### Visualize the data

Created a faceted histogram to compare the distribution of sizes between purple and red urchins. 

```{r}
ggplot(data = urchins_data, aes(x= SIZE))+
  geom_histogram(color= "black", fill= "white")+
  facet_grid(COMMON_NAME ~ .)+
  labs(x = "Size (cm)",
       y = "Count",
       title = "Distibution of sizes between purple and red urchins")
```

