---
title: 'Task 3: Santa Barbara urchins'
author: "Anna Zauner"
date: "10/23/2021"
output: html_document
---

## Overview

This task utilizes data from 5 sites along the Santa Barbara coast to compare the size distirbution of red and purple urchins using data visualization and summary statistics.....

**Data citation**: Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b. 


## Set up

```{r setup, include=TRUE, warning= FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# Attach packages
library(tidyverse)
library(dplyr)
library(janitor)
library(here)
library(kableExtra)
library(lubridate)

```

## Read in data

Read in with `read_csv` and cleaned names with `clean_names()`
```{r}
urchins_data <- read_csv(here("data", "urchins.csv")) %>% 
clean_names() %>% 
rename(size_cm = size)

```

# Section 1

Size comparisons, purple and red urchins. 

### Visualize the data

Created a faceted histogram to compare the distribution of sizes between purple and red urchins. 

```{r}
ggplot(data = urchins_data, aes(x= size_cm))+
  geom_histogram(color= "black", fill= "white", bins = 10)+
  facet_wrap(~common_name)+
  labs(x = "Size in diameter (cm)",
       y = "Count",
       title = "Distibution of sizes between purple and red urchins (ZAUNER)",
       caption = "Figure 1: Comparing the distribution of sizes in diameter (cm) between the purple and red urchins off the coast of Santa Barbara.")+
  theme(plot.caption = element_text(hjust = 0.5, size = 8))
```

## Descriptive statistics

Created a table to summarize statistic for purple and red urchins by finding mean, median, standard deviation and sample size. 

```{r}
urchins_summary <- urchins_data %>% 
  group_by(common_name) %>% 
  summarize(mean = mean(size_cm),
            median = median(size_cm),
            sd = sd(size_cm),
            sample_size = n()) %>% 
  kable(caption = "Table 1: Summary statistics of urchin sizes",
        col.names = c("Urchin name",
                      "Mean (cm)",
                      "Median (cm)",
                      "Standard Deviation",
                      "Sample Size"),
        digits = 1) %>% 
  kable_minimal()

urchins_summary
```


**Table 1 Takeaways:** The purple urchin population has a lower central tendency (4.5 cm mean, 4.5 cm median) than the red urchin population (6.8 cm mean, 7.0 cm median). The purple urchins are more concentrated around the central tendency, with a smaller spread and lower standard deviation (1.2) while red urchins are more spread with a higher standard deviation (1.9).


# Section 2

## Explore the purple urchin data

Compared purple urchin counts by site and year. Done by `filter()` only purple urchins, changing date class to "Date",and creating new columns. 

```{r}
purple_urchins <- urchins_data %>% 
  filter(common_name == "Purple Urchin") %>% 
  mutate(date_new = mdy(date)) %>% 
  mutate(year_new = year(date_new)) %>% 
  mutate(site_full = case_when(
    site == "NAPL" ~ "Naples Reef",
    site == "AQUE" ~ "Arroyo Quemado",
    site == "CARP" ~ "Carpinteria Reef",
    site == "MOHK" ~ "Mohawk Reef",
    site == "IVEE" ~ "Isla Vista Reef"
  ))

```
## Descriptive statistics

Created a table to store of total number of purple urchins observed by site and year. 

```{r}
total_purple_site_year <- purple_urchins %>%
  group_by(site_full, year_new) %>% 
  summarize(sample_size = n())
  
total_purple_site_year

```

## Data visualization

Created graph of counts of purple urchins recorded at each site over time.

```{r}

ggplot(data = total_purple_site_year, aes(x = year_new, y = sample_size))+
  geom_line(stat = "identity")+
  aes(color = site_full)+
  labs(x = "Year",
       y = "Total purple urchin count",
       title = "Total purple urchin count per site (2008-2020) (ZAUNER)",
       fill = "Site name",
       caption = "Figure 2: Counts of purple urchins recorded at each site along the coast of Santa Barbara from 2008 to 2020.")+
  theme(plot.caption = element_text(hjust = 0.5, size = 8))
  
  
```

**Figure 2 Takeaways:** 